Adaptive broadcast tree for multicores

### 2013-06-19 paper.tex
UMP breakdown is a section
Generating _two_ header files now, one with the actual C variable representing the model, the other with some meta data. Stupid me! Should have done this from the start. j;agkwl go;iqw
New reference. Separate pdf for visualizations.

% What problem are you solving, and why is it hard?
The problem is complexity and diversity of modern multicore
machines. Multicores come in many flavors depending on vendor and
generation of the machine. All of these machines have fundamentally
different characteristics such as number of nodes, NUMA
characteristics, interconnect topologies, propagation time on
interconnect and use of shared caches. These characteristics are hard
to understand, but important to consider for application performance.

@@

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Implementation}

@@

@@

\begin{itemize}
\item Other topologies (mainly based on~\cite{Bruck1992}):
  \begin{itemize}
  \item Fibonacci trees: more efficient since nodes further up the
    tree send more messages, but not considering the hierarchy of the
    machine 
  \item \nary trees: play with the out-degree of nodes. The ideal
    value for the out-degree depends on the machine. For
    $t_{propagation} << max(t_{send}, t_{receive})$, $n = 2$ should be
    optimal, since it still increases parallelism, but the latency
    induced by the $n$ send operation is not too high.
  \end{itemize}
\item Think about synchronous implementations and see if they are
  applicable for some of our machines.
\item Arguing that group communication/atomic broadcast is important
  and showing that we can do it well might be enough (appeal to atomic
  broadcasts as the foundation for other distributed algorithms, and
  the MPI primitives, that can/need to be mapped to
  group-communication (have a list of them)
  \begin{itemize}
  \item MPI\_Bcast: send a broadcast message to every node (including
    the sending one)
  \item MPI\_Scatter
  \item MPI\_Gather
  \item MPI\_Reduce
  \end{itemize}
  These primitives are synchronous. They return only once every other
  node received the message. 
\item Need higher-level applications, such as:
  \begin{itemize}
  \item capability system? \stefan{Ask Simon how expensive revocation
      currently is}
  \item a database with replication and consistency maintenance
    \begin{itemize}
    \item SharedDB? (buffers between operators, but no synchronization
      except for access to buffers, which are multiple writers, one
      reader)
    \item Crescando (some kind of state machine replication)
    \end{itemize}
  \end{itemize}
\item MPI collectives (as in~\cite{Tu2008})
\item \emph{Load balancing in a tree} We could build several rings,
  that do not always include all coordinators and work with read- and
  write-sets. We can select these rings such that they nicely
  integrate with the physical topology. They can also have different
  sizes (e.g.\ a smaller one for read operations and larger ones for
  writes).
\item \emph{Aggregation for convergecast} %
  Convergecast is a way of collecting information from nodes (i.e.\ it
  is kind of the reverse of a broadcast). On the return path,
  information can be aggregated. One example is batching, where several
  pieces of information are grouped in the same message to reduce packet
  processing overhead. Another example is aggregation, i.e.\
  pre-processing of data. One example is pre-calculating the average
  value of child messages if the sink node is not interested in
  individual values. %
  Similarly to what has been done in wireless sensor networks (where it
  also matters to reduce the number of messages, but for other reasons:
  power consumption), we can do aggregation in nodes. In difference to
  traditional distributed systems, this works, because it is easy to
  deploy custom software on every node in the network. Classical
  distributed systems do not typically allow this. Furthermore, reducing
  the number of messages at the price of higher complexity does not make
  sense in classical systems. %
  Examples for aggregation: number of nodes agreeing to something, find
  capabilities (concatenate core ids). 
\end{itemize}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{UMP performance}

@@

\newpage
\input{measurements/gruyere_topologies}
\input{measurements/ziger1_topologies}
\input{measurements/sbrinz1_topologies}

\input{graphs/visall}

>>>>

% What problem are you solving, and why is it hard?
The problem is complexity and diversity of modern multicore
machines. Multicores come in many flavors depending on vendor and
generation of the machine. All of these machines have fundamentally
different characteristics such as number of nodes, NUMA
characteristics, interconnect topologies, propagation time on
interconnect and use of shared caches. These characteristics are hard~\cite{Cavage2013}
to understand, but important to consider for application performance.

@@

\begin{figure}[htb]
  \centering
  \includegraphics[width=\linewidth]{Overview_Model_Evaluation_Parser}
  \caption{Overview of configuration of atomic broadcast using a Model 
    \stefan{Draw figure properly}}
  \label{fig:overview}
\end{figure}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Implementation}
\label{sec:implementation}

@@

\stefan{Describe benchmark application} 
\begin{itemize}
\item One process on every core
\item Management process on core 2 (i.e.\ not core 0)
\item Uses atomic broadcast as described in Section~\ref{sec:implementation}
\end{itemize}

@@

\begin{itemize}
\item Keywords: decomposition~\cite{Cavage2013}
\item Other topologies (mainly based on~\cite{Bruck1992}):
  \begin{itemize}
  \item Fibonacci trees: more efficient since nodes further up the
    tree send more messages, but not considering the hierarchy of the
    machine 
  \item \nary trees: play with the out-degree of nodes. The ideal
    value for the out-degree depends on the machine. For
    $t_{propagation} << max(t_{send}, t_{receive})$, $n = 2$ should be
    optimal, since it still increases parallelism, but the latency
    induced by the $n$ send operation is not too high.
  \end{itemize}
\item Think about synchronous implementations and see if they are
  applicable for some of our machines.
\item Arguing that group communication/atomic broadcast is important
  and showing that we can do it well might be enough (appeal to atomic
  broadcasts as the foundation for other distributed algorithms, and
  the MPI primitives, that can/need to be mapped to
  group-communication (have a list of them)
  \begin{itemize}
  \item MPI\_Bcast: send a broadcast message to every node (including
    the sending one)
  \item MPI\_Scatter
  \item MPI\_Gather
  \item MPI\_Reduce
  \end{itemize}
  These primitives are synchronous. They return only once every other
  node received the message. 
\item Need higher-level applications, such as:
  \begin{itemize}
  \item capability system? \stefan{Ask Simon how expensive revocation
      currently is}
  \item a database with replication and consistency maintenance
    \begin{itemize}
    \item SharedDB? (buffers between operators, but no synchronization
      except for access to buffers, which are multiple writers, one
      reader)
    \item Crescando (some kind of state machine replication)
    \end{itemize}
  \end{itemize}
\item MPI collectives (as in~\cite{Tu2008})
\item \emph{Load balancing in a tree} We could build several rings,
  that do not always include all coordinators and work with read- and
  write-sets. We can select these rings such that they nicely
  integrate with the physical topology. They can also have different
  sizes (e.g.\ a smaller one for read operations and larger ones for
  writes).
\item \emph{Aggregation for convergecast} %
  Convergecast is a way of collecting information from nodes (i.e.\ it
  is kind of the reverse of a broadcast). On the return path,
  information can be aggregated. One example is batching, where several
  pieces of information are grouped in the same message to reduce packet
  processing overhead. Another example is aggregation, i.e.\
  pre-processing of data. One example is pre-calculating the average
  value of child messages if the sink node is not interested in
  individual values. %
  Similarly to what has been done in wireless sensor networks (where it
  also matters to reduce the number of messages, but for other reasons:
  power consumption), we can do aggregation in nodes. In difference to
  traditional distributed systems, this works, because it is easy to
  deploy custom software on every node in the network. Classical
  distributed systems do not typically allow this. Furthermore, reducing
  the number of messages at the price of higher complexity does not make
  sense in classical systems. %
  Examples for aggregation: number of nodes agreeing to something, find
  capabilities (concatenate core ids). 
\end{itemize}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{UMP performance}

@@

\newpage
\input{measurements/gruyere_topologies}
\input{measurements/ziger1_topologies}
\input{measurements/sbrinz1_topologies}
\input{measurements/appenzeller_topologies}

%\input{graphs/visall}
